\section*{Acknowledgments}
\addcontentsline{toc}{section}{Acknowledgments}
First and foremost I am extremely grateful to my supervisors Laurent Romary and Benoît Sagot, who at the beginning of my Ph.D. accepted to supervise me even though I was coming from a different background and had absolutely no experience in Natural Language Processing (NLP) or Digital Humanities (DH). You both had the patience to teach me and accompany me throughout this Ph.D. journey, supporting me not only academically but also personally; you're both my mentors and for that I am extremely grateful. To Benoît, thank you so much for always being there during all the difficult deadlines ready to correct and advice. To Laurent, thank you so much for all your invaluable advise which undeniably helped go through the difficult situations that I encountered during this Ph.D.

I would like to thank all the members of the committee for agreeing to evaluate this thesis and its defense, for their comments, advice and questions which have greatly contributed to the improvement and finalization of this work. Special thanks to Anna Korhonen and Holger Schwenk, for accepting to read my manuscript in such a short span of time and for the helpful comments in your pre-reports. I would also like to thank the members of my \emph{comité de suivi de thèse}, for their support and their advice for the end of the thesis.

I would also like to thank the ANR BASNUM project (ANR-18-CE38-0003) for founding my Ph.D. thesis, as well as its members Geoffrey Williams, Ioana Galleron, Mohamed Khemakhem and Clarissa Stincone, with whom I had the pleasure of collaborating at the beginning of this Ph.D. thesis.

This thesis is not only the product of my work, but the fruit of numerous collaborations without which it would have never been possible. For my collaborators and coauthors, I would like to thank all the members of the \emph{CamemBERT} project, specially to Louis and Benjamin. We managed to publish a valuable and influential resource for French NLP and even for NLP in languages other than English. I will always be grateful to have participated in this project.

I also want to mention Rachel, Philippe and Alexandre from the \emph{Early Modern French} project; Tian Tian and Gaël Lejeune from the \emph{CLEF-HIPE2020} shared task; and Benoît Crabbé from the \emph{BERTrade} project. You taught me many things during our time collaborating together, and I cannot say enough how lucky I was to have the opportunity to work with you and how grateful I am to all of you.

To Simon Gabay, with whom I collaborated during the last part of my Ph.D. thesis, I'd like to thank for providing me with so much high quality data. Most of the experiments at the end of this manuscript would have never been possible without your work, so for this you have my whole gratitude.

To all the people at Masakhane and all the coauthors of the \emph{Quality at Glance} paper, specially to Julia and Isaac; I truly have no words to express how lucky I was to work with such an amazing group of people, you taught me about the importance of looking at my own data, and you also helped me restart the OSCAR project. Thank you so much for working with me and expressing interest in my corpus.

As the OSCAR project became larger, I no longer had the means to continue working on it alone, this is why I would like to express my deepest gratitude to Julien. Thank you so much for your patience, for taking the time to go through my messy code and making it readable, for managing our users while I was away and for taking care of OSCAR like it was your own. I only have gratitude and respect for you, you're truly one of the greatest engineers that I know, and even though we have sadly never met in person, know that you became a valuable friend and colleague to me during this last year.

To all my other colleagues at the ALMAnaCH team that I have worked with such as Djamé, Eric, Syrielle, Kim and Meriem; and to the newcomers such as Hugo, Rua and Roman. Thank you for making of Inria such a nice place to work at.

At the beginning of this Ph.D. thesis I was just a mathematician that did not have any clue about NLP or Machine Learning. However, I encountered a group of people that believed in me and supported me in working at Inria and starting this Ph.D. thesis in NLP. For this I would like to thank in particular Patrice, Alba and Professor Isis Truck.

While this Ph.D. thesis was an academic exercise, I cannot overstate how valuable was the personal support that the friends I made here in Paris gave me during the last four years. You made me feel at home in France and you actually helped me become one of your own; you accompanied me during very turbulent personal times and even through the pandemic. Some of you even became coauthors and collaborators. Thank you so much Yoann and Ariane, Alix and Jen, Clementine, Loïc, Murielle, Manon, Emilia, Ganesh, Tanti and Arij. I do not have words to truly express to you how thankful I am.

To all my Colombian friends that have always supported me, and specially to Zubieta, Juan Rafael, Pedro Pablo, Isabel, Susana, Henao, Pablo, Andrés, Maritza Carlos and Min. Thank you so much for sending me all your support from afar.

To my family, specially to my father, my grandmother, to Elena and Gloria and to my little Family in the US, Fercho and Beatriz. Thank you so much for all your support and your love, in spite of the distance and the pandemic keeping us apart for years, you were always there when I needed it the most. Thank you so much for always believing in me.

Finally, I'd like to thank Mathilde and her family, Paul, Anne, Thierry and Anne-Marie. Since the beginning you always welcomed me with open arms, and you have supported and accompanied me during the last two years. I have no words to express how grateful I am for everything that you have done for me and for the hospitality that you have always offered me. To Mathilde, you have all my gratitude for your constant support and specially for putting up with me during the last days of this Ph.D. thesis, I know it was difficult, but I do believe good times are ahead of us.
